---
title: "aha paper"
author: "Claudia"
date: '2023-02-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Predicting the impact of EHR implementation
This study explores the relationships between EHR implementation and health outcomes, including number of readmissions, total performance score, overall hospital rating, patient knowledge, quality of care, and care coordination. 

```{r}
#set working directory
getwd()
setwd("C:/Users/claud/OneDrive/Desktop/RA")
```

```{r}
#Import datasets
aha <- read.csv('./data/ehr_aha_cms_provider_id_cleaned_lean.csv',header = TRUE, sep = ",")
readmissions<- read.csv('./data/READMISSION REDUCTION_ehr.csv',header = TRUE, sep = ",")
tps<- read.csv('./data/hvbp_tps_10_28_2015_ehr.csv',header = TRUE, sep = ",")
rating<- read.csv('./data/HCAHPS - Hospital_ehr.csv',header = TRUE, sep = ",")
knowledge<- read.csv('./data/hvbp_imm2_11_10_2016.csv',na.string="Not Available",header = TRUE, sep = ",")
qoc<-read.csv('./data/hvbp_hcahps_11_10_2016.csv',na.string="Not Available",header = TRUE, sep = ",")
coordination<-read.csv('./data/HOSPITAL_QUARTERLY_QUALITYMEASURE_IPFQR_HOSPITAL.csv',na.string="Not Available",header = TRUE, sep = ",")
```

```{r}
View(aha)
colnames(aha)
```
```{r}
head(aha)
```

```{r}
str(aha)
```

```{r}
View(readmissions)
colnames(readmissions)
```
```{r}
head(readmissions)
```

```{r}
str(readmissions)
```

```{r}
nrow(readmissions)
ncol(readmissions)
```


```{r}
nrow(aha)
ncol(aha)
```

```{r}
nrow(tps)
ncol(tps)
```

```{r}
head(tps)
```

```{r}
str(tps)
```

```{r}
nrow(rating)
ncol(rating)
```

```{r}
str(rating)
```

```{r}
nrow(knowledge)
ncol(knowledge)
```
```{r}
str(knowledge)
```


```{r}
str(coordination)
```
```{r}
nrow(coordination)
ncol(coordination)
```

```{r}
str(qoc)
```
```{r}
nrow(qoc)
ncol(qoc)
```

```{r}
#check for NA's in aha dataset
colSums(is.na(aha))
```

```{r}
#check for NA's in readmissions dataset
colSums(is.na(readmissions))
```
```{r}
#check for NA's in tps dataset
colSums(is.na(tps))
```
```{r}
#check for NA's in ratings dataset
colSums(is.na(rating))
```
```{r}
#check for NA's in knowledge dataset
colSums(is.na(knowledge))
```


```{r}
#check for NA's in qoc dataset
colSums(is.na(qoc))
```
```{r}
#check for NA's in coordination dataset
colSums(is.na(coordination))
```
##Cleaning data
```{r}
#subset datasets with useful columns
knowledge<- knowledge[, c(1,8:15)]
coordination<- coordination[, c(1,109,130)]
qoc<- qoc[, c(1,28,36,44,60)]
```


```{r}
nrow(knowledge)
ncol(knowledge)
```


```{r}
nrow(coordination)
ncol(coordination)
```

```{r}
nrow(qoc)
ncol(qoc)
```
```{r}
head(knowledge)
```

```{r}
head(coordination)
```

```{r}
head(qoc)
```

```{r}
#renaming column names
colnames(knowledge)[colnames(knowledge) == "Provider_Number"] <- "Provider_ID"
colnames(qoc)[colnames(qoc) == "Provider.Number"] <- "Provider_ID"
colnames(coordination)[colnames(coordination) == "Provider_Number"] <- "Provider_ID"
```


##Merging datasets
```{r}
#merge datasets using Provider_ID
d1 <- rating
d2 <- readmissions
m1 <- merge(d1, d2 ,by="Provider_ID")
View(m1)
d3 <- tps
m2 <- merge(m1, d3, by="Provider_ID")
View(m2)
d4 <- aha
m3 <- merge(m2, d4,by="Provider_ID" )
View(m3)
d5<-knowledge
m4<-merge(d5, m3, by="Provider_ID")
View(m4)
d6<-qoc
m5<-merge(d6, m4 ,by="Provider_ID")
View(m5)
d7<-coordination
total<-merge(d7, m5, by="Provider_ID")
View(total)
```

##Final dataset
```{r}
colnames(total)
```

```{r}
total
```


```{r}
nrow(total)
ncol(total)
```
```{r}
str(total)
```

```{r}
#change int variables to numeric
total[sapply(total, is.integer)] <- lapply(total[sapply(total, is.integer)],as.numeric)
```

```{r}
str(total)
```
```{r}
#change chr variables to factor
total[sapply(total, is.character)] <- lapply(total[sapply(total, is.character)],as.factor)
```

```{r}
str(total)
```

```{r}
colSums(is.na(total))
```

```{r}
nrow(total)
ncol(total)
```

```{r}
#subset of dataset with no NA's
total_sub<-total[!is.na(total$HBIPS.7_65_Over_._of_Total)&!is.na(total$HBIPS.6_65_Over_._of_Total) &!is.na (total$NETWRK),]
```

```{r}
table(total_sub$NETWRK)
```

```{r}
table(total_sub$HBIPS.6_65_Over_._of_Total)
```

```{r}
table(total_sub$HBIPS.7_65_Over_._of_Total)
```

```{r}
sum(is.na(total_sub))
```

```{r}
nrow(total_sub)
ncol(total_sub)
```


##Linear regression
```{r}
#readmission
lm_readmission_1<- lm(readdm_total ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total_sub) 
```

```{r}
summary(lm_readmission_1)
```

```{r}
lm_readmission_2<- lm(readdm_total ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total) 
```

```{r}
summary(lm_readmission_2)
```
```{r}
lm_readmission_3<- lm(readdm_total ~ HOSPBD , data = total_sub)
```

```{r}
summary(lm_readmission_3)
```

```{r}
#ehr and size on readmission
lm_readmission_4<- lm(readdm_total ~ EHLTH * HOSPBD, data = total) 
```

```{r}
summary(lm_readmission_4)
```
```{r}
lm_readmission_5<- lm(readdm_total ~ EHLTH * HOSPBD, data = total_sub)
```

```{r}
summary(lm_readmission_5)
```

```{r}
#total performance score(TPS)
lm_tps_1 <- lm(TPS ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total_sub)
```

```{r}
summary(lm_tps_1)
```

```{r}
lm_tps_2<- lm(TPS ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total)
```

```{r}
summary(lm_tps_2)
```
```{r}
lm_tps_3 <- lm(TPS ~ location_metro , data = total_sub) 
```

```{r}
summary(lm_tps_3)
```
```{r}
#ehr and location metro on TPS
lm_tps_4<- lm(TPS ~ EHLTH * location_metro  , data = total)
```

```{r}
summary(lm_tps_4)
```
```{r}
lm_tps_5<- lm(TPS ~ EHLTH * location_metro  , data = total_sub)
```

```{r}
summary(lm_tps_5)
```

```{r}
#overall hospital rating
lm_rating_1 <- lm(HCAHPS_Linear_Mean_Value ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total_sub) 
```

```{r}
summary(lm_rating_1)
```

```{r}
lm_rating_2 <- lm(HCAHPS_Linear_Mean_Value ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total)
```

```{r}
summary(lm_rating_2)
```
```{r}
lm_rating_3 <- lm(HCAHPS_Linear_Mean_Value ~location_rural  , data = total_sub) 
```

```{r}
summary(lm_rating_3)
```

```{r}
#knowledge
lm_knowledge_1<-lm(IMM.2.Performance.Rate ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total_sub) 

```

```{r}
summary(lm_knowledge_1)
```

```{r}
lm_knowledge_2<-lm(IMM.2.Performance.Rate ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total)
```

```{r}
summary(lm_knowledge_2)
```
```{r}
lm_knowledge_3<-lm(IMM.2.Performance.Rate ~ HOSPBD + location_rural, data = total)
```

```{r}
summary(lm_knowledge_3)
```

```{r}
#quality of care
lm_qoc_1<-lm(Communication.about.Medicines.Performance.Rate ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total_sub)
```

```{r}
summary(lm_qoc_1)
```

```{r}
lm_qoc_2<-lm(Communication.about.Medicines.Performance.Rate ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total)
```

```{r}
summary(lm_qoc_2)
```
```{r}
lm_qoc_3<-lm(Communication.about.Medicines.Performance.Rate ~ location_rural , data = total)
```

```{r}
summary(lm_qoc_3)
```

```{r}
#care coordination
lm_coordination_1<-lm(HBIPS.7_65_Over_._of_Total ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total_sub)
```

```{r}
summary(lm_coordination_1)
```

```{r}
lm_coordination_2<-lm(HBIPS.7_65_Over_._of_Total ~ EHLTH + HOSPBD + location_metro + location_rural + control_nongov_nfp + control_investerowned_fp , data = total)
```

```{r}
summary(lm_coordination_2)
```
##Multicolinearity
```{r}
#multicolinearity 
#install.packages('car')
library(car)
vif(lm_coordination_1)
vif(lm_coordination_2)
```
```{r}
vif(lm_knowledge_1)
vif(lm_knowledge_2)
vif(lm_knowledge_3)
```

```{r}
vif(lm_qoc_1)
vif(lm_qoc_2)
```

```{r}
vif(lm_rating_1)
vif(lm_rating_2)
```

```{r}
vif(lm_readmission_1)
vif(lm_readmission_2)
```

```{r}
vif(lm_tps_1)
vif(lm_tps_2)
```


##Conclusion
I excluded columns in the datasets that were not useful and merged AHA's 2015-2016 dataset with data from CMS using "Provider_ID". After merging, I was left was 192 rows and 34 columns. I created a subset of the final dataset that had no missing values, which had 118 rows and 34 columns. The linear regression models showed counterintuitive results for readmission and TPS. 